Binocular combination in the autonomic nervous system

### 2025-01-14 manuscript.tex
Added high resolution supplementary figure

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Binocular combination in the autonomic nervous system},
  colorlinks=true,
  linkcolor=black,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{caption} \usepackage{setspace}\doublespacing \usepackage{lineno} \linenumbers
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

Binocular combination has been extensively studied in visual perception, where it is mediated by neurons in primary visual cortex (Hubel and Wiesel, \protect\hyperlink{ref-Hubel1962}{1962}). For pattern vision, binocular summation occurs at threshold, such that a lower contrast is required to detect a target shown to both eyes than a target shown to one eye (Campbell and Green, \protect\hyperlink{ref-Campbell1965}{1965}; Baker et al., \protect\hyperlink{ref-Baker2018}{2018}). At higher contrasts, a process of `ocularity invariance' is observed, in which the response to monocularly- and binocularly-presented patterns is equal (Baker et al., \protect\hyperlink{ref-Baker2007}{2007}; Moradi and Heeger, \protect\hyperlink{ref-Moradi2009}{2009}). This is explained by a process of interocular suppression that cancels out the additional excitatory drive caused by stimulating two eyes. Our recent work (Segala et al., \protect\hyperlink{ref-Segala2023}{2023}) showed that cortical signal combination for luminance flicker is substantially more linear than for spatial patterns, whereas there is evidence of interocular suppression in the pupil pathway. Here we measure the amplitude of pupil modulations in response to flickering stimuli directed at the luminance, L-M cone, S-cone, and melanopsin pathways (see Figure \ref{fig:spectraplots}e-l). Our key comparisons are between monocular and binocular stimulation, and in a dichoptic masking condition where the two eyes are stimulated at different frequencies. The results are interpreted using a contemporary model of binocular vision (Meese et al., \protect\hyperlink{ref-Meese2006}{2006}) implemented within a hierarchical Bayesian framework.

@@

To present synchronised stimulus modulations independently to each eye, two light engines (SpectraTuneLAB: LEDMOTIVE Technologies, LLC, Barcelona, Spain), each with 10 independently addressable LED colour channels, were integrated into a customised binocular viewing system. The light engines were operated via a Python interface to their REST API (Martin et al., \protect\hyperlink{ref-Martin2022}{2022}), which supports synchronous launch and playback of spectral sequences prepared in advance and stored in JSON format. When preparing the spectral sequences, the age of participants was used to account for the yellowing of the lenses. We used the silent substitution technique (Estévez and Spekreijse, \protect\hyperlink{ref-Estevez1982}{1982}) to selectively stimulate specific photoreceptor classes. Silent substitution exploits the fact that each photoreceptor class has a distinct spectral tuning that overlaps with the others. Using a multiprimary system, in which the primaries (i.e.~LEDs) have different spectra, it is possible to target one class of photoreceptors while maintaining the others at a constant activity level, effectively silencing them (Shapiro et al., \protect\hyperlink{ref-Shapiro1996}{1996}; Spitschan and Woelders, \protect\hyperlink{ref-Spitschan2018}{2018}, this paper also offers a clear explanation of how to implement silent substitution). We calculated silent substitution solutions using the \emph{PySilSub} toolbox (Martin et al., \protect\hyperlink{ref-Martin2023}{2023}), using linear algebra. The outputs of the two light engines (see Figure \ref{fig:spectraplots}b,c) were calibrated using an Ocean Optics Jaz spectroradiometer, which was wavelength-calibrated to an Argon lamp and intensity calibrated using a NIST-traceable light source. We used the Stockman and Sharpe (\protect\hyperlink{ref-Stockman2000}{2000}) 10-degree cone fundamentals, and estimates of melanopsin absorbance spectra from CIE S 026 (discussed in Martin et al., \protect\hyperlink{ref-Martin2023}{2023}) to calculate \(\alpha\)-opic irradiance.

@@

The diffuser discs were positioned in the objective planes of the lenses of a modified VR headset (SHINECON SC-G01, Dongguan Shinecon Industrial Co.~Ltd., Guangdong, China), which was used by the participants to view the stimuli. The stimuli were two discs of flickering light with a diameter of approximately 30 degrees, which were fused together into a cyclopean percept resembling a donut-shaped ring of light, similar to that used in other studies (e.g., Barrionuevo and Cao, \protect\hyperlink{ref-Barrionuevo2016}{2016}; Spitschan et al., \protect\hyperlink{ref-Spitschan2014}{2014}, \protect\hyperlink{ref-Spitschan2015}{2015}; Murray et al., \protect\hyperlink{ref-Murray2018}{2018}; Zele et al., \protect\hyperlink{ref-Zele2018}{2018}). The VR headset modifications allowed for small adjustments to account for individual differences in interpupillary distance and focal length. The use of this set up allowed us to modulate the stimuli in three different ocular configurations, similar to the ones we used in our previous study (Segala et al., \protect\hyperlink{ref-Segala2023}{2023}): monocular, binocular and dichoptic. In the monocular configuration, the unstimulated eye still saw a non-flickering disc of mean luminance. A schematic of the stimulation system is shown in Figure \ref{fig:spectraplots}a. Pupillometry data were collected using a binocular Pupil Core eye-tracker headset (Pupil Labs GmbH, Berlin, Germany, Kassner et al., \protect\hyperlink{ref-Kassner2014}{2014}) running at 120 Hz, and the signals were recorded with the Pupil Capture software.

@@

In the luminance experiment, the stimulus intensity was increased and decreased relative to the background, which we expected to modulate all photoreceptor classes (see Figure \ref{fig:spectraplots}e,i). In the L-M cone modulation experiment, we used silent substution to increase the L-cone activity, and simultaneously decrease the M-cone activity, during the first half-cycle of the sine wave. In the second half-cycle the polarity of the modulation reversed (see Figure \ref{fig:spectraplots}f,j). The maximum available L-M contrast was approximately 10\%. In the S-cone modulation experiment, we increased and decreased S-cone-directed signals, whilst keeping activity in the other photoreceptors constant (see Figure \ref{fig:spectraplots}g,k). Our system allowed a maximum contrast of 45\%. Finally, in the melanopsin experiment, we modulated the activity of the melanopsin-containing intrinsically photoreceptive retinal ganglion cells, whilst keeping cone activity constant (see Figure \ref{fig:spectraplots}h,l). The maximum available melanopsin contrast was 22\%. We assume that the activity of rods was constant at the high background luminance intensity used here, and so did not attempt to silence rod activity in any condition, as this would have greatly reduced the available dynamic range. Splatter on nominally silenced photoreceptors was very small (see Figure \ref{fig:spectraplots}j-l), well below the levels that would be expected to generate measureable pupil modulations. We also estimated activation of penumbral L and M cones (Spitschan et al., \protect\hyperlink{ref-Spitschan2015}{2015}; Barrionuevo and Cao, \protect\hyperlink{ref-Barrionuevo2016}{2016}; Zele et al., \protect\hyperlink{ref-Zele2019}{2019}), which was minimal (\(\le 1.5%
\) contrast; less than splatter on the open-field cones) for melanopsin-directed stimuli. We note that the temporal frequency of our modulation (0.5Hz) is well below the range where penumbral cone activation can elicit visible percepts, and that such percepts fade after around 1 second (Spitschan et al., \protect\hyperlink{ref-Spitschan2015}{2015}), and do not typically affect pupil responses (Spitschan et al., \protect\hyperlink{ref-Spitschan2014}{2014}).

@@

To describe our data, we used the same model described in our previous study (Segala et al., \protect\hyperlink{ref-Segala2023}{2023}). The model has the same general form as the first stage of the contrast gain control model proposed by Meese et al. (\protect\hyperlink{ref-Meese2006}{2006}), but omits the second stage. For the previous model that we used (Segala et al., \protect\hyperlink{ref-Segala2023}{2023}), the exponent of the numerator and denominator had fixed values of 2 and (implicitly) 1. Here, we allow these parameters (called \emph{p} and \emph{q}) to be free, in order to permit different shapes of contrast response function, e.g.~accelerating or saturating. The responses of the left eye and right eye channels are as follows:

@@

To estimate the extent of interocular suppression for each photoreceptor pathway, we fitted each data set using a Bayesian hierarchical implementation of a simple binocular combination model (Meese et al., \protect\hyperlink{ref-Meese2006}{2006}; Segala et al., \protect\hyperlink{ref-Segala2023}{2023}). Our primary objective was to compare posterior distributions of the weight of interocular suppression, which are shown in Figure \ref{fig:FirstHarmonicPlots}e. Consistent with our earlier observations, the strongest suppressive weight corresponds to the L-M and Melanopsin conditions, and the weakest suppression corresponds to the luminance and S-cone conditions, with virtually no overlap between the posterior distributions for weak and strong suppression. The model fits were of good quality, as shown by the curves in Figure \ref{fig:FirstHarmonicPlots}a-d,f-i. The fitted model parameters are given in Table \ref{tab:paramtable}.

@@

Finally, we inspected the response phase of our four stimulation conditions, given previous reports that these differ across pathways (Spitschan et al., \protect\hyperlink{ref-Spitschan2014}{2014}), and may be in antiphase for melanopsin and S-cone signals. Figure \ref{fig:phaseplots} shows the phase angles for the first (a) and second (b) harmonic frequencies for binocular stimulation (monocular stimulation produced very similar results). At the first harmonic, melanopsin and S-cone signals differed in phase by more than 90 degrees, though they were not fully in antiphase. The luminance and L-M responses were approximately in antiphase to each other, however this is likely due to our choice to modulate L+M- in the first half-cycle of the sine wave (corresponding to a luminance increase in the luminance condition), and L-M+ in the second half-cycle (corresponding to a luminance decrease in the luminance condition). Had we reversed this phase arrangement, we would likely have seen a close phase correspondence between the L-M and luminance conditions (another way to think about this is that L-cone decreases and M-cone increases are processed like luminance increases). Phase differences between the luminance, S-cone and melanopsin conditions are likely attributable to different lags in phototransduction at the earliest stage (i.e.~in the retina). At the second harmonic frequency (Figure \ref{fig:phaseplots}b), the luminance condition was again out of phase with the other three, and was approximately in quadrature phase with the L-M condition, and in antiphase with both the S-cone and melanopsin conditions (which were in phase with each other).

@@

Our recent psychophysical work (Baker et al., \protect\hyperlink{ref-Baker2024}{2024}) has looked at binocular interactions in the L-M and S-cone pathways, and compared these to the luminance pathway. Using a contrast discrimination paradigm with spatial modulations of luminance and colour, Baker et al. (\protect\hyperlink{ref-Baker2024}{2024}) found equally strong interocular suppression in all three pathways. This is rather different from our pupillometry results here, which demonstrate weaker suppression in the luminance and S-cone pathways than in the L-M pathway (see Figure \ref{fig:FirstHarmonicPlots}). However, the current experiments involve temporal modulations, which are quite distinct from the spatial modulations used in Baker et al. (\protect\hyperlink{ref-Baker2024}{2024}). Our other recent work (Segala et al., \protect\hyperlink{ref-Segala2023}{2023}) has shown that temporal luminance modulations involve much weaker interocular suppression in the cortical response than do spatial luminance modulations (Baker and Wade, \protect\hyperlink{ref-Baker2017}{2017}). These different normalization processes might reflect different priorities for spatial and temporal vision. Spatial vision aims to fuse images to provide binocular single vision, and benefits from `ocularity invariance' (Baker et al., \protect\hyperlink{ref-Baker2007}{2007}), in which visual appearance is constant when viewed with one eye or two. Temporal vision is critical for motion perception, which can involve alternative binocular computations, such as calculating velocity differences between the eyes (Kaestner et al., \protect\hyperlink{ref-Kaestner2019}{2019}). Of course the present measurements were of pupil size, which may be subject to different anatomical and functional constraints from those in the cortex. Future research should aim to extend our current findings to perception using psychophysical approaches.

@@

We have demonstrated that binocular combination of temporal flickering light in the autonomic nervous system depends on the photoreceptor pathway. We were able to elicit pupil responses by stimulating the periphery of the retina and we were able to record contrast response functions for all photoreceptor pathways. While all pathways showed non-linear combination, they varied in how the signals are combined, particularly in the weight of interocular suppression. This was strong (\(\omega \ge 1\)) for L-M and melanopsin signals at the first harmonic, and all pathways at the second harmonic. Suppression was weaker (\(\omega < 1\)) in the luminance pathway (consistent with previous work), and also for S-cone directed modulations.

@@

\leavevmode\hypertarget{ref-Baker2024}{}%
Baker DH, Hansford KJ, Segala FG, Morsi AY, Huxley RJ, Martin JT, Rockman M, Wade AR (2024) Binocular integration of chromatic and luminance signals. Journal of Vision Under review.

@@

@@

@@

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Binocular combination in the autonomic nervous system},
  colorlinks=true,
  linkcolor=black,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{caption} \usepackage{setspace}\doublespacing \usepackage{lineno} \linenumbers \newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

Binocular combination has been extensively studied in visual perception, where it is mediated by neurons in primary visual cortex (Hubel and Wiesel, \protect\hyperlink{ref-Hubel1962}{1962}). For pattern vision, binocular summation occurs at threshold, such that a lower contrast is required to detect a target shown to both eyes than a target shown to one eye (Campbell and Green, \protect\hyperlink{ref-Campbell1965}{1965}; Baker et al., \protect\hyperlink{ref-Baker2018}{2018}). At higher contrasts, the phenomenon of `ocularity invariance' is observed, in which the response to monocularly- and binocularly-presented patterns is equal (Baker et al., \protect\hyperlink{ref-Baker2007}{2007}; Moradi and Heeger, \protect\hyperlink{ref-Moradi2009}{2009}). This is explained by a process of interocular suppression that cancels out the additional excitatory drive caused by stimulating two eyes. Our recent work (Segala et al., \protect\hyperlink{ref-Segala2023}{2023}) showed that cortical signal combination for luminance flicker is substantially more linear than for spatial patterns, whereas there is evidence of interocular suppression in the pupil pathway. Here we measure the amplitude of pupil modulations in response to flickering stimuli presented as light flux, or directed towards the L-M cone, S-cone, and melanopsin pathways (see Figure \ref{fig:spectraplots}e-l). Our key comparisons are between monocular and binocular stimulation, and in a dichoptic masking condition where the two eyes are stimulated at different frequencies. The results are interpreted using a contemporary model of binocular vision (Meese et al., \protect\hyperlink{ref-Meese2006}{2006}) implemented within a hierarchical Bayesian framework.

@@

To present synchronised stimulus modulations independently to each eye, two light engines (SpectraTuneLAB, which are thermally stable across time with active cooling according to the manufacturers: LEDMOTIVE Technologies, LLC, Barcelona, Spain), each with 10 independently addressable LED colour channels, were integrated into a customised binocular viewing system. The light engines were operated via a Python interface to their REST API (Martin et al., \protect\hyperlink{ref-Martin2022}{2022}), which supports synchronous launch and playback of spectral sequences prepared in advance and stored in JSON format. A special command was commissioned from LEDMOTIVE to allow two different stimulus files to be launched simultaneously from the two devices. The synchronisation of the spectra from the two devices was tested and showed that they were synchronised to within approximately 3ms.

When preparing the spectral sequences, the age of participants was used to account for the yellowing of the lenses. We used the silent substitution technique (Estévez and Spekreijse, \protect\hyperlink{ref-Estevez1982}{1982}) to selectively stimulate specific photoreceptor classes. Silent substitution exploits the fact that each photoreceptor class has a distinct spectral tuning that overlaps with the others. Using a multiprimary system, in which the primaries (i.e.~LEDs) have different spectra, it is possible to target one class of photoreceptors while maintaining the others at a constant activity level, effectively silencing them (Shapiro et al., \protect\hyperlink{ref-Shapiro1996}{1996}; Spitschan and Woelders, \protect\hyperlink{ref-Spitschan2018}{2018}, this paper also offers a clear explanation of how to implement silent substitution). We calculated silent substitution solutions using the \emph{PySilSub} toolbox (Martin et al., \protect\hyperlink{ref-Martin2023}{2023}), using linear algebra. The outputs of the two light engines (see Figure \ref{fig:spectraplots}b,c) were calibrated using an Ocean Optics Jaz spectroradiometer, which was wavelength-calibrated to an Argon lamp and intensity calibrated using a NIST-traceable light source. Each primary was also linearized using a polynomial fit (see Figure \ref{fig:primarylinearity} for details). We used the Stockman and Sharpe (\protect\hyperlink{ref-Stockman2000}{2000}) 10-degree cone fundamentals, and estimates of melanopsin absorbance spectra from CIE S 026 (discussed in Martin et al., \protect\hyperlink{ref-Martin2023}{2023}) to calculate \(\alpha\)-opic irradiance.

@@

The diffuser discs were positioned in the objective planes of the lenses of a modified VR headset (SHINECON SC-G01, Dongguan Shinecon Industrial Co.~Ltd., Guangdong, China), which was used by the participants to view the stimuli. The stimuli were two discs of flickering light with a diameter of approximately 30 degrees, which were fused together into a cyclopean percept resembling a donut-shaped ring of light, similar to that used in other studies (e.g., Barrionuevo and Cao, \protect\hyperlink{ref-Barrionuevo2016}{2016}; Spitschan et al., \protect\hyperlink{ref-Spitschan2014}{2014}, \protect\hyperlink{ref-Spitschan2015}{2015}; Murray et al., \protect\hyperlink{ref-Murray2018}{2018}; Zele et al., \protect\hyperlink{ref-Zele2018}{2018}). The VR headset modifications allowed for small adjustments to account for individual differences in interpupillary distance and focal length. The use of this set up allowed us to modulate the stimuli in three different ocular configurations, similar to the ones we used in our previous study (Segala et al., \protect\hyperlink{ref-Segala2023}{2023}): monocular, binocular and dichoptic. In the monocular configuration, the unstimulated eye still saw a non-flickering disc of mean light flux. A schematic of the stimulation system is shown in Figure \ref{fig:spectraplots}a. Pupillometry data were collected using a binocular Pupil Core eye-tracker headset (Pupil Labs GmbH, Berlin, Germany, Kassner et al., \protect\hyperlink{ref-Kassner2014}{2014}) running at 120 Hz, and the signals were recorded with the Pupil Capture software.

@@

Silent substitution stimuli were prepared and calibrated for each participant with custom Python software (Martin et al., \protect\hyperlink{ref-Martin2023}{2023}) and Python scripts. Estimates of photoreceptor spectral sensitivities for each participant were constructed from the known photopigment absorbance spectra (Stockman and Sharpe, \protect\hyperlink{ref-Stockman2000}{2000}), taking account of the peak axial density of the respective photopigments, as well as lens (Pokorny et al., \protect\hyperlink{ref-Pokorny1987}{1987}; Stockman et al., \protect\hyperlink{ref-Stockman1999}{1999}; Stockman and Sharpe, \protect\hyperlink{ref-Stockman2000}{2000}) and macular pigment density (Bone et al., \protect\hyperlink{ref-Bone1988}{1988}; Stockman et al., \protect\hyperlink{ref-Stockman1999}{1999}), in accordance with the field size and age-dependant CIEPO06 observer model (CIE, \protect\hyperlink{ref-CIE2006}{2006}). The melanopic and rhodopic action spectra of the 32-year-old standard observer were taken from CIE S 026 (CIE, \protect\hyperlink{ref-CIE2018}{2018}) and then adjusted for age-related lens transmittance with a spectral correction function, in line with the standard. Macular pigment correction was not applied to the rhodopic and melanopic action spectra because rods are not present at the fovea and ipRGCs sit above the retinal pigment layer (Trieschmann et al., \protect\hyperlink{ref-Trieschmann2007}{2007}).

In the light flux experiment, the stimulus intensity was increased and decreased relative to the background, which we expected to modulate all photoreceptor classes (see Figure \ref{fig:spectraplots}e,i). In the L-M cone modulation experiment, we used silent substitution to increase the L-cone activity, and simultaneously decrease the M-cone activity, during the first half-cycle of the sine wave. In the second half-cycle the polarity of the modulation reversed (see Figure \ref{fig:spectraplots}f,j). The maximum available L-M contrast was approximately 10\%. In the S-cone modulation experiment, we increased and decreased S-cone-directed signals, whilst keeping activity in the other photoreceptors constant (see Figure \ref{fig:spectraplots}g,k). Our system allowed a maximum contrast of 45\%. Finally, in the melanopsin experiment, we modulated the activity of the melanopsin-containing intrinsically photoreceptive retinal ganglion cells, whilst keeping cone activity constant (see Figure \ref{fig:spectraplots}h,l). The maximum available melanopsin contrast was 22\%. We assume that the activity of rods was constant at the high background luminance intensity used here, and so did not attempt to silence rod activity in any condition, as this would have greatly reduced the available dynamic range. Splatter on nominally silenced photoreceptors was very small (see Figure \ref{fig:spectraplots}j-l), well below the levels that would be expected to generate measurable pupil modulations, although we can observe that the rods may not be saturated in the L-M and melanopsin-directed stimuli (Figure \ref{fig:spectraplots}j and \ref{fig:spectraplots}l) and could intrude in these two experiments. We also estimated activation of penumbral L and M cones (Spitschan et al., \protect\hyperlink{ref-Spitschan2015}{2015}; Barrionuevo and Cao, \protect\hyperlink{ref-Barrionuevo2016}{2016}; Zele et al., \protect\hyperlink{ref-Zele2019}{2019}), which was minimal (\(\le 1.5%
\) contrast; less than splatter on the open-field cones) for melanopsin-directed stimuli. We note that the temporal frequency of our modulation (0.5Hz) is well below the range where penumbral cone activation can elicit visible percepts, and that such percepts fade after around 1 second (Spitschan et al., \protect\hyperlink{ref-Spitschan2015}{2015}), and do not typically affect pupil responses (Spitschan et al., \protect\hyperlink{ref-Spitschan2014}{2014}).

@@

To quantitatively summarise our data, we used the same model described in our previous study (Segala et al., \protect\hyperlink{ref-Segala2023}{2023}). The model has the same general form as the first stage of the contrast gain control model proposed by Meese et al. (\protect\hyperlink{ref-Meese2006}{2006}), but omits the second stage. For the previous model that we used (Segala et al., \protect\hyperlink{ref-Segala2023}{2023}), the exponent of the numerator and denominator had fixed values of 2 and (implicitly) 1. Here, we allow these parameters (called \emph{p} and \emph{q}) to be free, in order to permit different shapes of contrast response function, e.g.~accelerating or saturating. The responses of the left eye and right eye channels are as follows:

@@

To estimate the extent of interocular suppression for each photoreceptor pathway, we fitted each data set using a Bayesian hierarchical implementation of a simple binocular combination model (Meese et al., \protect\hyperlink{ref-Meese2006}{2006}; Segala et al., \protect\hyperlink{ref-Segala2023}{2023}). Our primary objective was to compare posterior distributions of the weight of interocular suppression, which are shown in Figure \ref{fig:FirstHarmonicPlots}e. Consistent with our earlier observations, the strongest suppressive weight corresponds to the L-M and Melanopsin conditions, and the weakest suppression corresponds to the light flux and S-cone conditions, with virtually no overlap between the posterior distributions for weak and strong suppression. The model fits were of good quality, as shown by the curves in Figure \ref{fig:FirstHarmonicPlots}a-d,f-i. The fitted model parameters are given in Table \ref{tab:paramtable}.

@@

Finally, we inspected the response phase of our four stimulation conditions, given previous reports that these differ across pathways (Spitschan et al., \protect\hyperlink{ref-Spitschan2014}{2014}), and may be in antiphase for melanopsin and S-cone signals. Figure \ref{fig:phaseplots} shows the phase angles for the first (a) and second (b) harmonic frequencies for binocular stimulation (monocular stimulation produced very similar results). At the first harmonic, melanopsin and S-cone signals differed in phase by more than 90 degrees, though they were not fully in antiphase. The light flux and L-M responses were approximately in antiphase to each other, however this is likely due to our choice to modulate L+M- in the first half-cycle of the sine wave (corresponding to a luminance increase in the light flux condition), and L-M+ in the second half-cycle (corresponding to a luminance decrease in the light flux condition). Had we reversed this phase arrangement, we would likely have seen a close phase correspondence between the L-M and light flux conditions (another way to think about this is that L-cone decreases and M-cone increases are processed like luminance increases). Phase differences between the light flux, S-cone and melanopsin conditions are likely attributable to different lags in phototransduction at the earliest stage (i.e.~in the retina). At the second harmonic frequency (Figure \ref{fig:phaseplots}b), the light flux condition was again out of phase with the other three, and was approximately in quadrature phase with the L-M condition, and in antiphase with both the S-cone and melanopsin conditions (which were in phase with each other). We note that the marked phase differences between conditions make the possibility that our results are dominated by rod activity relatively unlikely (e.g.~in the L-M and melanopsin conditions, which have very different phase profiles).

@@

Our recent psychophysical work (Baker et al., \protect\hyperlink{ref-Baker2024}{2024}) has looked at binocular interactions in the L-M and S-cone pathways, and compared these to the light flux pathway. Using a contrast discrimination paradigm with spatial modulations of light flux and colour, Baker et al. (\protect\hyperlink{ref-Baker2024}{2024}) found equally strong interocular suppression in all three pathways. This is rather different from our pupillometry results here, which demonstrate weaker suppression in the light flux and S-cone pathways than in the L-M pathway (see Figure \ref{fig:FirstHarmonicPlots}). However, the current experiments involve temporal modulations, which are quite distinct from the spatial modulations used in Baker et al. (\protect\hyperlink{ref-Baker2024}{2024}). Our other recent work (Segala et al., \protect\hyperlink{ref-Segala2023}{2023}) has shown that temporal luminance modulations involve much weaker interocular suppression in the cortical response than do spatial luminance modulations (Baker and Wade, \protect\hyperlink{ref-Baker2017}{2017}). These different normalization processes might reflect different priorities for spatial and temporal vision. Spatial vision aims to fuse images to provide binocular single vision, and benefits from `ocularity invariance' (Baker et al., \protect\hyperlink{ref-Baker2007}{2007}), in which visual appearance is constant when viewed with one eye or two. Temporal vision is critical for motion perception, which can involve alternative binocular computations, such as calculating velocity differences between the eyes (Kaestner et al., \protect\hyperlink{ref-Kaestner2019}{2019}). Of course the present measurements were of pupil size, which may be subject to different anatomical and functional constraints from those in the cortex. Future research should aim to extend our current findings to perception using psychophysical approaches.

@@

We have demonstrated that binocular combination of temporal flickering light in the autonomic nervous system depends on the photoreceptor pathway stimulated. We were able to elicit pupil responses by stimulating the periphery of the retina and we were able to record contrast response functions for all photoreceptor pathways. While all pathways showed non-linear combination, they varied in how the signals are combined, particularly in the weight of interocular suppression. This was strong (\(\omega \ge 1\)) for L-M and melanopsin signals at the first harmonic, and all pathways at the second harmonic. Suppression was weaker (\(\omega < 1\)) in the light flux pathway (consistent with previous work), and also for S-cone directed modulations.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

Supported by Biotechnology and Biological Sciences Research Council grant BB/V007580/1 awarded to DHB and ARW.

\hypertarget{declaration-of-interests}{%
\section{Declaration of interests}\label{declaration-of-interests}}

The authors declare no competing interests.

@@

\leavevmode\hypertarget{ref-Baker2024}{}%
Baker DH, Hansford KJ, Segala FG, Morsi AY, Huxley RJ, Martin JT, Rockman M, Wade AR (2024) Binocular integration of chromatic and luminance signals. J Vis 24(12):7.

@@

\leavevmode\hypertarget{ref-Bone1988}{}%
Bone RA, Landrum JT, Fernandez L, Tarsis SL (1988) Analysis of the macular pigment by HPLC: Retinal distribution and age study. Investigative Ophthalmology \& Visual Science 29:843--849.

@@

\leavevmode\hypertarget{ref-CIE2006}{}%
CIE (2006) Fundamental chromaticity diagram with physiological axes. CIE Central Bureau. Available at: \url{https://cie.co.at/publications/fundamental-chromaticity-diagram-physiological-axes-part-1}.

\leavevmode\hypertarget{ref-CIE2018}{}%
CIE (2018) CIE system for metrology of optical radiation for ipRGC-influenced responses to light. CIE Central Bureau. Available at: \url{https://doi.org/10.25039/S026.2018}.

@@

\leavevmode\hypertarget{ref-Pokorny1987}{}%
Pokorny J, Smith VC, Lutze M (1987) Aging of the human lens. Applied Optics 26:1437 Available at: \url{http://dx.doi.org/10.1364/AO.26.001437}.

@@

\hypertarget{supplementary-material}{%
\section{Supplementary material}\label{supplementary-material}}
